---
title: "Beyance and Taylor Swift"
author: "Hunter Kempf"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Music to my Ears: An analysis of Taylor Swift and Beyance Albums

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggwordcloud)
library(tidytext)
```

Disclaimer: 
I found some issues in the sales and charts csv files. Namely both were missing Beyonce's Lemonade album. It appears to be an error in the wikipedia parsing script featured on the [github repo](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-29/readme.md) as of Sept 28, 2020 


```{r}
beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
```

```{r}
# used regex101 to build a pattern that worked to extract out the dates
sales <- sales %>% mutate(released_date = mdy(sub("^([A-z]* [0-9]*, [0-9]*).*", "\\1", released)))
ggplot(
  data = sales %>% filter(country %in% c("WW","World")),
  aes(x= released_date,y = sales/1e6, color=artist)
) + geom_line() + 
  geom_point() + 
  ylab("Sales in Millions of Copies") + 
  xlab("Album Release Date") +
  labs(title="Music to my Ears",
       subtitle="",
       color = "Artist",
       caption="Source :  | Analysis: @The_DataViz"
       ) + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold"))
```

## Taylor Swift Lyrics Analysis

### Attempting to use ggwordcloud or wordcloud2 but neither work for me

```{r}
set.seed(42)
ggplot(love_words_small, aes(label = word, size = speakers)) +
  geom_text_wordcloud_area(
    mask = png::readPNG(system.file("extdata/hearth.png",
      package = "ggwordcloud", mustWork = TRUE
    )),
    rm_outside = TRUE
  ) +
  scale_size_area(max_size = 40) +
  theme_minimal()
```

```{r}
library(wordcloud2) 
 
letterCloud( demoFreq, word = "R", color='random-light' , backgroundColor="black")

```

```{r}
taylor_swift_lyrics 
```

```{r message=FALSE, warning=FALSE}
title_in_song <- taylor_swift_lyrics %>% 
  mutate(Title2 = Title) %>%
  unnest_tokens(word,Title2) %>%
  group_by(Title) %>%
  summarize(title_word_length=n()) %>% 
  group_by(title_word_length) %>%
  summarize(count=n())
title_in_song
```

```{r}
word <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(word, Lyrics) %>% 
  group_by(Title,word) %>% 
  count() %>% 
  filter(tolower(Title)==word) %>%
  arrange(desc(n)) 
word
```

```{r}
bigram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(bigram, Lyrics, token = "ngrams", n = 2) %>% 
  group_by(Title,bigram) %>% 
  count() %>% 
  filter(tolower(Title)==bigram) %>%
  arrange(desc(n)) 
bigram
```

```{r}
trigram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(trigram, Lyrics, token = "ngrams", n = 3) %>% 
  group_by(Title,trigram) %>% 
  count() %>% 
  filter(tolower(Title)==trigram) %>%
  arrange(desc(n)) 
trigram
```

```{r}
quadgram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(quadgram, Lyrics, token = "ngrams", n = 4) %>% 
  group_by(Title,quadgram) %>% 
  count() %>% 
  filter(tolower(Title)==quadgram) %>%
  arrange(desc(n)) 
quadgram
```

```{r}
fivegram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(fivegram, Lyrics, token = "ngrams", n = 5) %>% 
  group_by(Title,fivegram) %>% 
  count() %>% 
  filter(tolower(Title)==fivegram) %>%
  arrange(desc(n)) 
fivegram
```

```{r}
sixgram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(sixgram, Lyrics, token = "ngrams", n = 6) %>% 
  group_by(Title,sixgram) %>% 
  count() %>% 
  filter(tolower(Title)==sixgram) %>%
  arrange(desc(n)) 
sixgram
```

```{r}
sevengram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(sevengram, Lyrics, token = "ngrams", n = 7) %>% 
  group_by(Title,sevengram) %>% 
  count() %>% 
  filter(tolower(Title)==sevengram) %>%
  arrange(desc(n)) 
sevengram
```

```{r}
eightgram <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(eightgram, Lyrics, token = "ngrams", n = 8) %>% 
  group_by(Title,eightgram) %>% 
  count() %>% 
  filter(tolower(Title)==eightgram) %>%
  arrange(desc(n)) 
eightgram
```

```{r}
title_in_song_count <- c(length(word$word),
  length(bigram$bigram),
  length(trigram$trigram),
  length(quadgram$quadgram),
  length(fivegram$fivegram),
  length(sixgram$sixgram),
  length(sevengram$sevengram),
  length(eightgram$eightgram))

title_in_song_plot <- title_in_song %>% 
  mutate(title_in_song_count = title_in_song_count,
         title_not_in_song = count-title_in_song_count) %>%
  select(title_word_length,title_in_song_count,title_not_in_song) %>%
  gather(key = "variable",value="count",-title_word_length)
title_in_song_plot
```

```{r}
ggplot(
  data = title_in_song_plot,
  aes(x= title_word_length,y=count,fill=variable)
) + geom_col(position="fill") + 
  ylab("% of Songs")
```

# Colors in Taylor Swift songs 

```{r}
color_counts <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(word, Lyrics) %>% 
  mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(Album,Title,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(Album,Title,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r}
border <- rep("white",12) 
border[c(4,11)] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 132 Taylor Swift Songs",
       caption="Source :  | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

# Colors in Beyance songs 

```{r}
color_counts <- beyonce_lyrics %>% 
  #mutate(Lyrics = str_replace_all(line,"\n","")) %>% 
  unnest_tokens(word, line) %>% 
  #mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(song_name,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(song_name,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r}
beyonce_lyrics %>% select(song_id) %>% unique() %>% nrow()
```


```{r}
border <- rep("white",15) 
border[6] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 390 Beyance Songs",
       caption="Source :  | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
