---
title: "Beyance and Taylor Swift"
author: "The DataViz"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Music to my Ears: An analysis of Taylor Swift and Beyance Albums

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggwordcloud)
library(tidytext)
```

Disclaimer: 
I found some issues in the sales and charts csv files. Namely both were missing Beyonce's Lemonade album. It appears to be an error in the wikipedia parsing script featured on the [github repo](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-29/readme.md) as of Sept 28, 2020 


```{r}
beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
```

```{r}
# used regex101 to build a pattern that worked to extract out the dates
sales <- sales %>% mutate(released_date = mdy(sub("^([A-z]* [0-9]*, [0-9]*).*", "\\1", released)))
ggplot(
  data = sales %>% filter(country %in% c("WW","World")),
  aes(x= released_date,y = sales/1e6, color=artist)
) + geom_line() + 
  geom_point() + 
  ylab("Sales in Millions of Copies") + 
  xlab("Album Release Date") +
  labs(title="Music to my Ears",
       subtitle="",
       color = "Artist",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday| Analysis: @The_DataViz"
       ) + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold"))
```

## Taylor Swift Lyrics Analysis

### Attempting to use ggwordcloud or wordcloud2 but neither work for me

```{r}
set.seed(42)
ggplot(love_words_small, aes(label = word, size = speakers)) +
  geom_text_wordcloud_area(
    mask = png::readPNG(system.file("extdata/hearth.png",
      package = "ggwordcloud", mustWork = TRUE
    )),
    rm_outside = TRUE
  ) +
  scale_size_area(max_size = 40) +
  theme_minimal()
```

```{r}
library(wordcloud2) 
 
letterCloud( demoFreq, word = "R", color='random-light' , backgroundColor="black")

```

```{r}
taylor_swift_lyrics
```

### Tswift Song Title in Lyrics

```{r message=FALSE, warning=FALSE}
title_in_song <- taylor_swift_lyrics %>% 
  mutate(Title2 = Title) %>%
  unnest_tokens(word,Title2) %>%
  group_by(Title) %>%
  summarize(title_word_length=n()) %>% 
  group_by(title_word_length) %>%
  summarize(count=n())
title_in_song
```

```{r}
tswift_lyrics_title_match <- function(title_lengths){
  print(paste("Initiating with:",title_lengths[1]))
  current_df <- taylor_swift_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[1]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[1]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    
  for (title_length in title_lengths[2:length(title_lengths)]) {
    print(paste("Looping for:",title_length))
    loop_df <- taylor_swift_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[title_length]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[title_length]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    current_df <- rbind(current_df,loop_df)
  }
  return(current_df)
}
lyric_in_tswift_song <- tswift_lyrics_title_match(title_in_song$title_word_length)
lyric_in_tswift_song %>% arrange(desc(n))
```

```{r}
title_in_song_count <- lyric_in_tswift_song %>%group_by(title_word_length) %>% count() %>% select(title_word_length,count_title_matches=n)
title_in_song_count
```


```{r}
title_in_song_plot <- title_in_song %>% 
  left_join(title_in_song_count,by="title_word_length") %>%
  mutate(count_title_matches = ifelse(is.na(count_title_matches),0,count_title_matches),
         True = count_title_matches,
         False = count-count_title_matches) %>%
  select(title_word_length,True,False) %>%
  gather(key = "variable",value="count",-title_word_length)
title_in_song_plot
```
```{r message=FALSE, warning=FALSE}
title_in_song_plot %>% 
  group_by(variable) %>% 
  summarize(total = sum(count)) %>% 
  spread(variable,total)%>%
  mutate(perc = True/(True+False)) %>%
  pull(perc)
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
ggplot(
  data = title_in_song_plot,
  aes(x= title_word_length,y=count,fill=variable)
) + geom_col(position="fill") + 
  ylab("% of Songs")+ 
  xlab("Word Count of Song Title") + 
  labs(title="Music to my Ears",
       fill="Title in Lyrics",
       subtitle="Taylor uses exact Lyrics as a Song Title 89% of the time",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("tswift_titles.png")
```

### Beyonce Song Title in Lyrics

As you can see the Beyonce Song titles are way more messy 

```{r}
beyonce_lyrics$song_name%>% unique()
```

In an effort to clean them up I will get rid of all text inside a () or [] which mostly ends up being featured artists, remixes etc this will help but stuff wont be perfect

```{r}
first<- sub("\\(.*\\)","",beyonce_lyrics$song_name%>% unique())
second <- sub("\\[.*\\]","",first%>% unique()) %>% str_trim()
second
```


```{r message=FALSE, warning=FALSE}
clean_title <- data.frame(song_name = beyonce_lyrics$song_name %>% unique())%>%
  mutate(mid =  sub("\\(.*\\)","",song_name),
         cleaner_title = str_trim(sub("\\[.*\\]","",mid))) 
unique_title <- data.frame(song_name =clean_title$cleaner_title %>% unique())
title_in_song <- unique_title %>%
  mutate(group = song_name) %>%
  unnest_tokens(word,group) %>%
  group_by(song_name) %>%
  summarize(title_word_length=n()) %>% 
  group_by(title_word_length) %>%
  summarize(count=n()) 
title_in_song
```
```{r}
beyonce_cleaner_lyrics$Title %>% unique() %>% length()
```


```{r}
beyonce_cleaner_lyrics %>%
  group_by(Title) %>% slice(which.max(nchar(as.character(Lyrics))))
```

```{r message=FALSE, warning=FALSE}
beyonce_cleaner_lyrics <- beyonce_lyrics %>% 
    group_by(song_id,song_name) %>% 
    summarize(Lyrics = paste0(line, collapse = " ")) %>% 
    mutate(mid =  sub("\\(.*\\)","",song_name),
         Title = str_trim(sub("\\[.*\\]","",mid))) %>%
  ungroup()%>%
  select(-c(song_id,mid)) %>% 
  group_by(Title) %>%
  slice(which.max(nchar(as.character(Lyrics)))) %>%
    select(Title,Lyrics)
```


```{r}
beyonce_lyrics_title_match <- function(title_lengths){
  print(paste("Initiating with:",title_lengths[1]))
  current_df <- beyonce_cleaner_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[1]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[1]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    
  for (title_length in title_lengths[2:length(title_lengths)]) {
    print(paste("Looping for:",title_length))
    loop_df <- beyonce_cleaner_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[title_length]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[title_length]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    current_df <- rbind(current_df,loop_df)
  }
  return(current_df)
}
lyric_in_beyonce_song <- beyonce_lyrics_title_match(title_in_song$title_word_length) %>% select(-words)
lyric_in_beyonce_song %>% arrange(desc(n))
```

```{r}
title_in_song_count <- lyric_in_beyonce_song %>%group_by(title_word_length) %>% count() %>% select(title_word_length,count_title_matches=n)
title_in_song_count
```


```{r}
title_in_song_plot <- title_in_song %>% 
  left_join(title_in_song_count,by="title_word_length") %>%
  mutate(count_title_matches = ifelse(is.na(count_title_matches),0,count_title_matches),
         True = count_title_matches,
         False = count-count_title_matches) %>%
  select(title_word_length,True,False) %>%
  gather(key = "variable",value="count",-title_word_length)
title_in_song_plot
```

```{r message=FALSE, warning=FALSE}
title_in_song_plot %>% 
  group_by(variable) %>% 
  summarize(total = sum(count)) %>% 
  spread(variable,total)%>%
  mutate(perc = True/(True+False)) %>%
  pull(perc)
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
ggplot(
  data = title_in_song_plot,
  aes(x= title_word_length,y=count,fill=variable)
) + geom_col(position="fill") + 
  ylab("% of Songs")+ 
  xlab("Word Count of Song Title") + 
  labs(title="Music to my Ears",
       fill="Title in Lyrics",
       subtitle="Beyonce uses exact Lyrics as a Song Title 64% of the time",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("beyonce_titles.png")
```

# Colors in Taylor Swift songs 

```{r}
color_counts <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n","")) %>% 
  unnest_tokens(word, Lyrics) %>% 
  mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(Album,Title,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(Album,Title,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
border <- rep("white",12) 
border[c(4,11)] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 132 Taylor Swift Songs",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("tswift_colors.png")
```

# Colors in Beyance songs 

```{r}
color_counts <- beyonce_lyrics %>% 
  #mutate(Lyrics = str_replace_all(line,"\n","")) %>% 
  unnest_tokens(word, line) %>% 
  #mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(song_name,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(song_name,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r}
beyonce_lyrics %>% select(song_id) %>% unique() %>% nrow()
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
border <- rep("white",15) 
border[6] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 390 Beyance Songs",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("beyance_colors.png")
```
