---
title: "Beyance and Taylor Swift"
author: "The DataViz"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Music to my Ears: An analysis of Taylor Swift and Beyance Albums

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggwordcloud)
library(tidytext)
library(usmap)
library(cowplot)
```


```{r}
beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
## Google Trends YouTube Search data: https://trends.google.com/trends/explore?date=all_2008&geo=US&gprop=youtube&q=%2Fm%2F0dl567,%2Fm%2F01mpq7s
timeline <- readr::read_csv("multiTimeline.csv",skip=2)
state_results <- readr::read_csv("geoMap.csv",skip=2)
```

```{r}
# used regex101 to build a pattern that worked to extract out the dates
sales <- sales %>% mutate(released_date = mdy(sub("^([A-z]* [0-9]*, [0-9]*).*", "\\1", released)))
ggplot(
  data = sales %>% filter(country %in% c("WW","World")),
  aes(x= released_date,y = sales/1e6, color=artist)
) + geom_line() + 
  geom_point() + 
  ylab("Sales in Millions of Copies") + 
  xlab("Album Release Date") +
  labs(title="Music to my Ears",
       subtitle="",
       color = "Artist",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday| Analysis: @The_DataViz"
       ) + 
  theme_bw() + 
  theme(plot.title = element_text(face="bold"))
```

## Taylor Swift Lyrics Analysis

### Attempting to use ggwordcloud or wordcloud2 but neither work for me

```{r}
set.seed(42)
ggplot(love_words_small, aes(label = word, size = speakers)) +
  geom_text_wordcloud_area(
    mask = png::readPNG(system.file("extdata/hearth.png",
      package = "ggwordcloud", mustWork = TRUE
    )),
    rm_outside = TRUE
  ) +
  scale_size_area(max_size = 40) +
  theme_minimal()
```

```{r}
library(wordcloud2) 
 
letterCloud( demoFreq, word = "R", color='random-light' , backgroundColor="black")

```

```{r}
taylor_swift_lyrics
```

### Tswift Song Title in Lyrics

```{r message=FALSE, warning=FALSE}
title_in_song <- taylor_swift_lyrics %>% 
  mutate(Title2 = Title) %>%
  unnest_tokens(word,Title2) %>%
  group_by(Title) %>%
  summarize(title_word_length=n()) %>% 
  group_by(title_word_length) %>%
  summarize(count=n())
title_in_song
```

```{r}
tswift_lyrics_title_match <- function(title_lengths){
  print(paste("Initiating with:",title_lengths[1]))
  current_df <- taylor_swift_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[1]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[1]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    
  for (title_length in title_lengths[2:length(title_lengths)]) {
    print(paste("Looping for:",title_length))
    loop_df <- taylor_swift_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[title_length]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[title_length]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    current_df <- rbind(current_df,loop_df)
  }
  return(current_df)
}
lyric_in_tswift_song <- tswift_lyrics_title_match(title_in_song$title_word_length)
lyric_in_tswift_song %>% arrange(desc(n))
```

```{r}
title_in_song_count <- lyric_in_tswift_song %>%group_by(title_word_length) %>% count() %>% select(title_word_length,count_title_matches=n)
title_in_song_count
```


```{r}
title_in_song_plot <- title_in_song %>% 
  left_join(title_in_song_count,by="title_word_length") %>%
  mutate(count_title_matches = ifelse(is.na(count_title_matches),0,count_title_matches),
         True = count_title_matches,
         False = count-count_title_matches) %>%
  select(title_word_length,True,False) %>%
  gather(key = "variable",value="count",-title_word_length)
title_in_song_plot
```
```{r message=FALSE, warning=FALSE}
title_in_song_plot %>% 
  group_by(variable) %>% 
  summarize(total = sum(count)) %>% 
  spread(variable,total)%>%
  mutate(perc = True/(True+False)) %>%
  pull(perc)
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
ggplot(
  data = title_in_song_plot,
  aes(x= title_word_length,y=count,fill=variable)
) + geom_col(position="fill") + 
  ylab("% of Songs")+ 
  xlab("Word Count of Song Title") + 
  labs(title="Music to my Ears",
       fill="Title in Lyrics",
       subtitle="Taylor uses exact Lyrics as a Song Title 89% of the time",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("tswift_titles.png")
```

### Beyonce Song Title in Lyrics

As you can see the Beyonce Song titles are way more messy 

```{r}
beyonce_lyrics$song_name%>% unique()
```

In an effort to clean them up I will get rid of all text inside a () or [] which mostly ends up being featured artists, remixes etc this will help but stuff wont be perfect

```{r}
first<- sub("\\(.*\\)","",beyonce_lyrics$song_name%>% unique())
second <- sub("\\[.*\\]","",first%>% unique()) %>% str_trim()
second
```


```{r message=FALSE, warning=FALSE}
clean_title <- data.frame(song_name = beyonce_lyrics$song_name %>% unique())%>%
  mutate(mid =  sub("\\(.*\\)","",song_name),
         cleaner_title = str_trim(sub("\\[.*\\]","",mid))) 
unique_title <- data.frame(song_name =clean_title$cleaner_title %>% unique())
title_in_song <- unique_title %>%
  mutate(group = song_name) %>%
  unnest_tokens(word,group) %>%
  group_by(song_name) %>%
  summarize(title_word_length=n()) %>% 
  group_by(title_word_length) %>%
  summarize(count=n()) 
title_in_song
```

Clean up format into one song per row similar to the tswift format also remove any data in () or [] most of which are features or remix titles. 
```{r message=FALSE, warning=FALSE}
beyonce_cleaner_lyrics <- beyonce_lyrics %>% 
    group_by(song_id,song_name) %>% 
    summarize(Lyrics = paste0(line, collapse = " ")) %>% 
    mutate(mid =  sub("\\(.*\\)","",song_name),
         Title = str_trim(sub("\\[.*\\]","",mid))) %>%
  ungroup()%>%
  select(-c(song_id,mid)) %>% 
  group_by(Title) %>%
  slice(which.max(nchar(as.character(Lyrics)))) %>%
    select(Title,Lyrics)
```

```{r}
beyonce_cleaner_lyrics$Title %>% unique() %>% length()
```

```{r}
beyonce_lyrics_title_match <- function(title_lengths){
  print(paste("Initiating with:",title_lengths[1]))
  current_df <- beyonce_cleaner_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[1]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[1]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    
  for (title_length in title_lengths[2:length(title_lengths)]) {
    print(paste("Looping for:",title_length))
    loop_df <- beyonce_cleaner_lyrics %>% 
                mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
                unnest_tokens(words, Lyrics, token = "ngrams", n = title_lengths[title_length]) %>% 
                group_by(Title,words) %>% 
                count() %>% 
                mutate(title_word_length=title_lengths[title_length]) %>%
                filter(tolower(Title)==words) %>%
                arrange(desc(n)) 
    current_df <- rbind(current_df,loop_df)
  }
  return(current_df)
}
lyric_in_beyonce_song <- beyonce_lyrics_title_match(title_in_song$title_word_length) %>% select(-words)
lyric_in_beyonce_song %>% arrange(desc(n))
```

```{r}
title_in_song_count <- lyric_in_beyonce_song %>%group_by(title_word_length) %>% count() %>% select(title_word_length,count_title_matches=n)
title_in_song_count
```


```{r}
title_in_song_plot <- title_in_song %>% 
  left_join(title_in_song_count,by="title_word_length") %>%
  mutate(count_title_matches = ifelse(is.na(count_title_matches),0,count_title_matches),
         True = count_title_matches,
         False = count-count_title_matches) %>%
  select(title_word_length,True,False) %>%
  gather(key = "variable",value="count",-title_word_length)
title_in_song_plot
```

```{r message=FALSE, warning=FALSE}
title_in_song_plot %>% 
  group_by(variable) %>% 
  summarize(total = sum(count)) %>% 
  spread(variable,total)%>%
  mutate(perc = True/(True+False)) %>%
  pull(perc)
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
ggplot(
  data = title_in_song_plot,
  aes(x= title_word_length,y=count,fill=variable)
) + geom_col(position="fill") + 
  ylab("% of Songs")+ 
  xlab("Word Count of Song Title") + 
  labs(title="Music to my Ears",
       fill="Title in Lyrics",
       subtitle="Beyonce uses exact Lyrics as a Song Title 64% of the time",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("beyonce_titles.png")
```

```{r}

```


# Colors in Taylor Swift songs 

```{r}
color_counts <- taylor_swift_lyrics %>% 
  mutate(Lyrics = str_replace_all(Lyrics,"\n"," ")) %>% 
  unnest_tokens(word, Lyrics) %>% 
  mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(Album,Title,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(Album,Title,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r}
color_counts %>% ungroup() %>% group_by(Album,Title) %>% summarize(sum=sum(n)) %>% arrange(desc(sum))
```

```{r}
color_counts %>% group_by(Title) %>% count() %>% arrange(desc(n))
```

```{r}
color_counts %>% filter(Title=="invisible string") %>% pull(Color)
```


```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
border <- rep("white",12) 
border[c(4,11)] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 132 Taylor Swift Songs",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("tswift_colors.png")
```

# Colors in Beyance songs 

```{r}
color_counts <- beyonce_lyrics %>% 
  #mutate(Lyrics = str_replace_all(line,"\n","")) %>% 
  unnest_tokens(word, line) %>% 
  #mutate(word = str_replace(word,"grey","gray"))%>%
  group_by(song_name,word) %>% 
  count()%>% 
  mutate(is_color = word %in% colors()) %>%
  filter(is_color==T) %>%
  select(song_name,Color=word,n)%>%
  arrange(desc(n))
color_counts
```

```{r}
color_counts %>% ungroup() %>% select(-Color) %>% group_by(song_name) %>% summarize(sum=sum(n)) %>% arrange(desc(sum))
```

```{r}
color_counts %>% group_by(song_name) %>% count() %>% arrange(desc(n))
```
```{r}
color_counts %>% filter(song_name=="Lemonade Film (Script)") %>% pull(Color)
```


```{r message=FALSE, warning=FALSE}
color_counts_Total <- color_counts %>% 
  group_by(Color) %>%
  summarize(count=sum(n)) %>%
  arrange(desc(count))
color_counts_Total
```

```{r}
beyonce_lyrics %>% select(song_id) %>% unique() %>% nrow()
```


```{r fig.height=4.05, fig.width=6.55, message=FALSE, warning=FALSE}
border <- rep("white",15) 
border[6] <- "black"
ggplot(
  data = color_counts_Total,
  aes(x = reorder(Color, desc(count)), y = count)
) + geom_col(fill = reorder(color_counts_Total$Color, 
                            desc(color_counts_Total$count)),
             color = border) +
  xlab("") + 
  ylab("Total Count") + 
  labs(title="Music to my Ears",
       subtitle="Colors Used in 390 Beyoncé Songs",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("beyonce_colors.png")
```



### Youtube Search Artist Timeline from Google Trends

```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
taylor_color = "#b31942"
beyonce_color = "#0a3161"
timeline <- timeline %>% mutate(Month_date = lubridate::parse_date_time(Month,"%Y-%m"))
ggplot(
  data=timeline,
  aes(x=Month_date)
) + geom_line(aes(y=`Taylor Swift: (United States)`),color=taylor_color,alpha=.7)+ 
  geom_line(aes(y=`Beyoncé: (United States)`),color=beyonce_color,alpha=.7) + 
  annotate(
    geom = "curve", x = ISOdate(2007, 6, 1), y = 87, xend = ISOdate(2008, 11, 1), yend = 50, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(
    geom = "curve", x = ISOdate(2008, 5, 1), y = 92, xend = ISOdate(2009, 8, 1), yend = 100, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2007, 1, 1), y = 90, size = 3,
           label = "Fearless (re-release)", 
           hjust = "left",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2010, 11, 1), y = 62, xend = ISOdate(2010, 11, 1), yend = 50, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2010, 11, 1), y = 65, size = 3,
           label = "Speak Now", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2012, 11, 1), y = 92, xend = ISOdate(2012, 10, 1), yend = 80, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2012, 11, 1), y = 95, size = 3,
           label = "Red", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2014, 12, 1), y = 92, xend = ISOdate(2014, 11, 1), yend = 83, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2014, 12, 1), y = 95, size = 3,
           label = "1989", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2017, 11, 1), y = 32, xend = ISOdate(2017, 11, 1), yend = 17, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2017, 11, 1), y = 35, size = 3,
           label = "Reputation", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2019, 9, 1), y = 22, xend = ISOdate(2019, 8, 1), yend = 13, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2019, 9, 1), y = 25, size = 3,
           label = "Lover", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2020, 7, 1), y = 22, xend = ISOdate(2020, 7, 1), yend = 9, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=taylor_color
  ) +
  annotate(geom = "text", x = ISOdate(2020, 7, 1), y = 25, size = 3,
           label = "Folklore", 
           hjust = "mid",color=taylor_color)+
  annotate(
    geom = "curve", x = ISOdate(2007, 12, 1), y = 52, xend = ISOdate(2008, 10, 1), yend = 70, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=beyonce_color
  ) +
  annotate(geom = "text", x = ISOdate(2006, 7, 1), y = 49, size = 3,
           label = "I Am... Sasha Fierce", 
           hjust = "left",color=beyonce_color)+
  annotate(
    geom = "curve", x = ISOdate(2011, 8, 1), y = 72, xend = ISOdate(2011, 7, 1), yend = 55, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=beyonce_color
  ) +
  annotate(geom = "text", x = ISOdate(2011, 7, 1), y = 75, size = 4,
           label = "4", 
           hjust = "left",color=beyonce_color)+
  annotate(
    geom = "curve", x = ISOdate(2013, 12, 1), y = 72, xend = ISOdate(2013, 12, 1), yend = 58, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=beyonce_color
  ) +
  annotate(geom = "text", x = ISOdate(2013, 12, 1), y = 75, size = 3,
           label = "Beyoncé", 
           hjust = "mid",color=beyonce_color)+
  annotate(
    geom = "curve", x = ISOdate(2016, 5, 1), y = 52, xend = ISOdate(2016, 4, 15), yend = 31, 
    curvature = .1, arrow = arrow(length = unit(2, "mm")),color=beyonce_color
  ) +
  annotate(geom = "text", x = ISOdate(2016, 5, 1), y = 55, size = 3,
           label = "Lemonade", 
           hjust = "mid",color=beyonce_color)+
  ylab("Relative Popularity")+
  xlab("")+ 
  labs(title="Music to my Ears",
       subtitle="Timeline of Youtube Relative Search History and Artist Album Releases (2008-2020)",
       caption="Source: Rosie Baillie and Dr. Sara Stoudt via #TidyTuesday and Google Trends | Analysis: @The_DataViz"
       ) +
  theme_bw()+ 
  theme(plot.title = element_text(face="bold"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggsave("Timeline.png")
```	

### Youtube Search US Map

```{r}
state_perc <- state_results %>% mutate(
  `Taylor Swift` = parse_number(`Taylor Swift: (1/1/08 - 9/30/20)`),
  `Beyoncé` = parse_number(`Beyoncé: (1/1/08 - 9/30/20)`)
) %>% 
  mutate(
    Artist = ifelse(`Taylor Swift`>`Beyoncé`,"Taylor Swift","Beyoncé"),
    Artist = factor(Artist,levels = c("Taylor Swift","Beyoncé")),
    Value = ifelse(`Taylor Swift`>`Beyoncé`,`Taylor Swift`,`Beyoncé`)/100
  )%>%
  select(state=Region,Artist,Value)
state_perc
```

```{r}
beyonce_states <- state_perc %>% filter(Artist=="Beyoncé")
tswift_states <- state_perc %>% filter(Artist=="Taylor Swift")
```

```{r message=FALSE, warning=FALSE}
artist_map <- plot_usmap(data = state_perc, values = "Artist",color="grey") + 
  scale_fill_manual(
    values = c(taylor_color,beyonce_color), name = "Artist"
  ) + theme(legend.position = "right",legend.title.align=0.5)
artist_map
```

```{r message=FALSE, warning=FALSE}
taylor_map <- plot_usmap(data = tswift_states, values = "Value", color = "grey",include = tswift_states$state) + 
  scale_fill_continuous(
    low = "#F0D1D9", high = taylor_color, name = "",
    limits = c(.5,.75),
    breaks = c(.50, .55, .60, .65, .70,.75),
    label = scales::percent_format(accuracy = 1)) + 
  theme(legend.position = "right",legend.title.align=0.5)
taylor_map
```

```{r message=FALSE, warning=FALSE}
beyonce_map <- plot_usmap(data = beyonce_states, values = "Value", color = "grey",include = beyonce_states$state) + 
  scale_fill_continuous(
    low = "#CED6DF", high = beyonce_color,name = "",
    limits = c(.5,.75),
    breaks = c(.50, .55, .60, .65, .70,.75),
    label = scales::percent_format(accuracy = 1)) + 
  theme(legend.position = "right",legend.title.align=0.5)
beyonce_map
```

```{r message=FALSE, warning=FALSE}
split<- plot_grid(taylor_map,beyonce_map, 
          labels = c("", ""),
          ncol = 2)
combo <- plot_grid(artist_map,split, 
          labels = c("", ""),
          ncol = 1) + 
  labs(title="Music to my Ears",
       subtitle="More Popular Artist in Youtube Relative Search History (2008-2020)",
       caption="Source: Google Trends | Analysis: @The_DataViz"
       ) 
combo
```

```{r}
10*4.51/7.29
```


```{r fig.height=6.19, fig.width=10, message=FALSE, warning=FALSE}
title_gg <- ggplot() + 
  labs(title = "Music to my Ears", subtitle = "More Popular Artist in Youtube Relative Search History (2008-2020)") + 
  theme_bw() + theme(plot.title = element_text(face="bold"),
                     panel.border = element_blank(),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
caption_gg <- ggplot() + 
  labs(caption="Source: Google Trends | Analysis: @The_DataViz") + 
  theme_bw() + 
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
plot_grid(title_gg, combo,caption_gg, ncol = 1, rel_heights = c(0.15, 1))
ggsave("youtube_map.png")
```

