---
title: "Great American Beer Awards"
author: "The DataViz"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggimage)
```

Tidy Tuesday Data from Great American Beer Festival
```{r}
beer_awards <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-20/beer_awards.csv')
```

### Additional Data from brewbound on craft breweries per state
https://www.brewbound.com/breweries?displayOutOfBiz=False
Data collected on 10/20/2020 by selecting the state and looking at the count on the map

```{r}
N <- 51
craftBrewers <- data.frame(state=rep("", N),craftBreweriesCount=rep(NA, N),
                 stringsAsFactors=FALSE)  
craftBrewers[1, ] <- list("AL",29)
craftBrewers[2, ] <- list("AK",32)
craftBrewers[3, ] <- list("AZ",82)
craftBrewers[4, ] <- list("AR",18)
craftBrewers[5, ] <- list("CA",665)
craftBrewers[6, ] <- list("CO",324)
craftBrewers[7, ] <- list("CT",59)
craftBrewers[8, ] <- list("DE",19)
craftBrewers[9, ] <- list("DC",13)
craftBrewers[10, ] <- list("FL",211)
craftBrewers[11, ] <- list("GA",66)
craftBrewers[12, ] <- list("HI",17)
craftBrewers[13, ] <- list("ID",45)
craftBrewers[14, ] <- list("IL",163)
craftBrewers[15, ] <- list("IN",108)
craftBrewers[16, ] <- list("IA",64)
craftBrewers[17, ] <- list("KS",27)
craftBrewers[18, ] <- list("KY",33)
craftBrewers[19, ] <- list("LA",32)
craftBrewers[20, ] <- list("ME",81)
craftBrewers[21, ] <- list("MD",70)
craftBrewers[22, ] <- list("MA",125)
craftBrewers[23, ] <- list("MI",214)
craftBrewers[24, ] <- list("MN",133)
craftBrewers[25, ] <- list("MS",12)
craftBrewers[26, ] <- list("MO",86)
craftBrewers[27, ] <- list("MT",64)
craftBrewers[28, ] <- list("NE",33)
craftBrewers[29, ] <- list("NV",40)
craftBrewers[30, ] <- list("NH",46)
craftBrewers[31, ] <- list("NJ",89)
craftBrewers[32, ] <- list("NM",53)
craftBrewers[33, ] <- list("NY",315)
craftBrewers[34, ] <- list("NC",198)
craftBrewers[35, ] <- list("ND",9)
craftBrewers[36, ] <- list("OH",155)
craftBrewers[37, ] <- list("OK",24)
craftBrewers[38, ] <- list("OR",217)
craftBrewers[39, ] <- list("PA",109)
craftBrewers[40, ] <- list("RI",16)
craftBrewers[41, ] <- list("SC",57)
craftBrewers[42, ] <- list("SD",15)
craftBrewers[43, ] <- list("TN",73)
craftBrewers[44, ] <- list("TX",218)
craftBrewers[45, ] <- list("UT",24)
craftBrewers[46, ] <- list("VT",44)
craftBrewers[47, ] <- list("VA",168)
craftBrewers[48, ] <- list("WA",291)
craftBrewers[49, ] <- list("WV",24)
craftBrewers[50, ] <- list("WI",143)
craftBrewers[51, ] <- list("WY",22)
```



### Weighting of Medals using NY Times Point system described in the article below
https://www.topendsports.com/events/summer/medal-tally/rankings-weighted.htm
TLDR: Gold   = 4 
      Silver = 2
      Bronze = 1

```{r}
beer_awards <- beer_awards %>% 
                  mutate(medal_NYTPoints = ifelse(medal=="Gold",4,ifelse(medal=="Silver",2,1)),
                         state = ifelse(state=="wa","WA",state),
                         state = ifelse(state=="Ak","AK",state))
```

## Group By Summaries

```{r}
statePoints <- beer_awards %>% group_by(state) %>% summarize(totalPoints = sum(medal_NYTPoints)) %>% 
                   ungroup() %>% arrange(state)
statePoints
```


```{r}
brewersAndPoints <- craftBrewers %>% left_join(statePoints, by="state") 
brewersAndPoints <- data.frame(state = state.abb,stateName = state.name) %>% inner_join(brewersAndPoints, by="state")
```

## West Virginia is the only state without a brewery with a top 3 finisher
```{r}
brewersAndPoints %>% filter(is.na(totalPoints))
```


```{r}
statesBybrewersAndPoints <- brewersAndPoints %>% mutate(  totalPoints = replace_na(totalPoints,0),
                              pointsPerCraftBrewer = totalPoints/craftBreweriesCount) %>% 
                 slice_max(pointsPerCraftBrewer,n=10)
statesBybrewersAndPoints
```
Add in Flag File Location to DF
```{r}
dir = "/Users/hunterkempf/Documents/GitHub/Tidy-Tuesday/"
img = list.files(paste0(dir,'10-20-2020/','Flags'), full.names = TRUE)
buttonImg = c()
squareImg = c()
for (imagePath in img){
  if (grepl("button", imagePath, fixed = TRUE)){
    buttonImg <- c(buttonImg,imagePath)
  }
  else{
    squareImg <- c(squareImg,imagePath)
  }
}
statesBybrewersAndPoints <- statesBybrewersAndPoints %>% arrange(state) %>% mutate(img = squareImg) %>% arrange(desc(pointsPerCraftBrewer)) %>% mutate(rank = 1:10) %>% 
                                mutate(stateName = reorder(stateName,rank))
```

Make graph of top 10 states by pointsPerCraftBrewer
Flags from VectorFlags.com

```{r}
statesBybrewersAndPoints %>% mutate(stateName = reorder(stateName,rank))
```
State Flag color scheme from www.schemecolor.com
```{r}
colors <- c("#002868", "#002868", "#76A1A0", "#002868", "#002A86", 
            "#0F204B", "#002986", "#002A86", "#B71234",  "#AB0635")
stripeColor <- c("#FFC324", "#FFFFFF", "#F0DC82", "#BF0A30", "#076A21", 
            "#FFB612", "#FFCC18", "#FFEA0F", "#BD8A5E",  "#FFFFFF")
ggplot(
  data = statesBybrewersAndPoints,
  aes(stateName,pointsPerCraftBrewer)
  ) + 
  ggpattern::geom_col_pattern(
    aes(pattern = "stripe", fill = stateName, pattern_fill = stripeColor),
    colour                   = 'black',
    pattern_density          = 0.35,
    pattern_key_scale_factor = 1.3) +
  geom_image(y = -.5,
             aes(image = img), 
             size = rep(0.09, 10),  # Play with this a bit
             by = "width") +
  scale_fill_manual(values = colors) +
  scale_color_manual(values = colors) +
  scale_y_continuous(labels = scales::comma) +
  expand_limits(y = -.6) +

  labs(x = "", y="Medal Points per Craft Brewery",
       title="Top 10 States By Medal Points per Craft Brewery",
       subtitle = 'Using New York Times Medal Weighted Point System',
       caption="Data Sources: Great American Beer Festival & brewbound.com\nFlag Image Source: VectorFlags.com | Analysis: @The_DataViz") +
  theme_bw()+
  theme(plot.title = element_text(face="bold"),
        legend.position = "none")
ggsave("pointsPerStatePerCraftBrewery.png")
```













